---
title: "Predictive Policing in Philadelphia"
author: "Frank Chen"
date: "2024-07-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(tidyverse)
library(tidycensus)
library(tigris)
library(sf)
library(ggplot2)
library(measurements)
library(scales)
library(extrafont)
library(colorspace)
library(viridis)
library(raster)
library(spdep)
library(spatstat)
library(tmap)
library(FNN)
library(grid)
library(gridExtra)
library(kableExtra)
library(classInt)

loadfonts(device = "win")
project_path <- "D:/research/predictive policing/data"
census_api_key("97557b2638bba3200febf4b8196a3444e9628887", overwrite = TRUE)
palette <- c("#c0513d", "#fdb724", "#1ca89d", "#a5bb49")
phl_crs <- "ESRI:102728"
```

```{r functions}
plot.theme <- function(){
  theme(text = element_text(family = "Calibri", size = 16), 
        plot.title = element_text(hjust = 0, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 14, margin = margin(r = 15, unit = "pt")),
        plot.caption = element_text(size = 12, vjust = -1.5),
        legend.position = "bottom",
        legend.margin = margin(t = 0, unit = "pt"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(color = "gray85"),
        panel.grid.minor.y = element_line(color = "gray95"))
}

map.theme <- function(base_size = 12, title_size = 16) {
  theme(
    text = element_text(color = "black"),
    plot.title = element_text(size = title_size,color = "black"),
    plot.subtitle=element_text(face = "italic"),
    plot.caption=element_text(hjust = 0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text.x = element_text(size = 14))
}

nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <-
    as.matrix(measureFrom)
  measureTo_Matrix <-
    as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
  output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}
```

## Introduction


## The Current Conditions of Narcotics Possession in Philadelphia

```{r read geo data}
neigh <- 
  st_read(paste0(
    project_path, "/philadelphia-neighborhoods.geojson"
    )) %>% 
  st_transform(crs = phl_crs)

phl_boundary <- 
  st_read(paste0(
    project_path, "/City_Limits.geojson"
    )) %>% 
  dplyr::select(geometry) %>% 
  st_transform(crs = phl_crs)

st_area(phl_boundary)
  
po_district <- 
  st_read(paste0(
    project_path, "/Police_District.geojson"
    )) %>% 
  st_transform(crs = phl_crs)

po_serv_area <- 
  st_read(paste0(
    project_path, "/Boundaries_PSA.geojson"
    )) %>% 
  st_transform(crs = phl_crs)

po_station <- 
  st_read(paste0(
    project_path, "/Police_Stations.geojson"
    )) %>% 
  st_transform(crs = phl_crs) %>% 
  dplyr::select(geometry) %>% 
  mutate(Legend = "Police_Station")
```

```{r read crime data}
crime_22 <- 
  read.csv(paste0(
    project_path, "/crime_incidents_2022.csv"
    ))

crime_23 <- 
  read.csv(paste0(
    project_path, "/crime_incidents_2023.csv"
    ))

felony_22 <- 
  crime_22 %>% 
  filter(ucr_general %in% c("100", "200", "300", "400", "500")) %>% 
  dplyr::select(Y = lat, X = lng) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(crs = phl_crs) %>%
  mutate(Legend = "felony")

felony_23 <- 
  crime_23 %>% 
  filter(ucr_general %in% c("100", "200", "300", "400", "500")) %>% 
  dplyr::select(Y = lat, X = lng) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(crs = phl_crs) %>%
  mutate(Legend = "felony")
```

```{r 311 data}
pub_ser <- read.csv(paste0(
    project_path, "/public_cases_311.csv"
    ))

abd_vehicle <- 
  pub_ser %>% 
  filter(service_name == "Abandoned Vehicle") %>% 
  dplyr::select(Y = lat, X = lon) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(crs = phl_crs) %>%
  mutate(Legend = "Abandoned_Cars")

dan_building <- 
  pub_ser %>% 
  filter(service_name == "Dangerous Building Complaint") %>% 
  dplyr::select(Y = lat, X = lon) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(crs = phl_crs) %>%
  mutate(Legend = "Dangerous_Building")

graffiti <- 
  pub_ser %>% 
  filter(service_name == "Graffiti Removal") %>% 
  dplyr::select(Y = lat, X = lon) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(crs = phl_crs) %>%
  mutate(Legend = "Graffiti")

sanitation <- 
  pub_ser %>% 
  filter(str_detect(service_name, "Sanitation")) %>% 
  dplyr::select(Y = lat, X = lon) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(crs = phl_crs) %>%
  mutate(Legend = "Sanitation")

st_light_out <- 
  pub_ser %>% 
  filter(service_name == "Street Light Outage") %>% 
  dplyr::select(Y = lat, X = lon) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(crs = phl_crs) %>%
  mutate(Legend = "Street_Light_Outage")

dan_park <- 
  pub_ser %>% 
  filter(service_name == "Parks and Rec Safety and Maintenance") %>% 
  dplyr::select(Y = lat, X = lon) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(crs = phl_crs) %>%
  mutate(Legend = "Unsafe_Parks")
```

```{r read liquor data}
bar <- 
  st_read(paste0(
    project_path, "/bars.geojson"
    )) %>% 
  st_transform(crs = phl_crs) %>% 
  dplyr::select(geometry) %>% 
  mutate(Legend = "Bar")

pubs <- 
  st_read(paste0(
    project_path, "/pubs.geojson"
    )) %>% 
  st_transform(crs = phl_crs) %>% 
  dplyr::select(geometry) %>% 
  mutate(Legend = "Pubs")

night_club <- 
  st_read(paste0(
    project_path, "/night_club.geojson"
    )) %>% 
  st_transform(crs = phl_crs) %>% 
  dplyr::select(geometry) %>% 
  mutate(Legend = "Night_club")

onsite_liquor <- 
  rbind(bar, pubs, night_club) %>% 
  mutate(Legend = "Onsite_Liquor")
  
liquor_retail <- 
  read.csv(paste0(
    project_path, "/liquor_retail.csv"
  )) %>% 
  st_as_sf(wkt = "geometry") %>% 
  st_set_crs(4326) %>% 
  dplyr::select(geometry) %>% 
  st_transform(crs = phl_crs) %>% 
  mutate(Legend = "Liquor_Retail")
```

```{r traffic data}
# https://www.dvrpc.org/traffic/
# Annual Average Daily Traffic (aadt)
traffic_count <- 
  st_read(paste0(
    project_path, "/Traffic_Count_Locations.geojson"
    )) %>% 
  filter(setyear == "2022",
         ! is.na(aadt)) %>% 
  st_transform(crs = phl_crs) %>% 
  dplyr::select(Traffic_Volume = aadt, geometry)
```

## Fishnet

Granularity: A 0.25 mile square provides a good balance between being granular enough to identify hotspots and large enough to be manageable in analysis.

Urban Fit: It aligns well with typical urban block sizes, covering roughly 4-6 city blocks, which is a practical size for police patrols and resource allocation.

Data Resolution: It allows for a detailed resolution that can capture variations in crime patterns effectively without overwhelming data analysts with too many small grids.

Given Philadelphia's block length of about 400-600 feet, a 0.25 mile (1,320 feet) square unit would encompass approximately 2-3 blocks in each direction

```{r makinig fishnet}
fishnet <- 
  st_make_grid(phl_boundary,
               cellsize = conv_unit(250, "m", "ft"), 
               square = TRUE) %>%
  .[phl_boundary] %>%            # fast way to select intersecting polygons
  st_sf() %>%
  mutate(uniqueID = 1:n())

crime_net <- 
  dplyr::select(felony_22) %>% 
  mutate(count_felony = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(count_felony = replace_na(count_felony, 0),
         uniqueID = 1:n(),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

ggplot() +
  geom_sf(data = crime_net, aes(fill = count_felony), color = NA) +
  scale_fill_viridis() +
  labs(title = "Count of felony for the fishnet") +
  map.theme()
```

```{r var net}
vars_net <- 
  rbind(abd_vehicle,dan_building,graffiti,
        sanitation, st_light_out, dan_park,
        onsite_liquor, liquor_retail, po_station) %>%
  st_join(., fishnet, join = st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
    full_join(fishnet) %>%
    spread(Legend, count, fill=0) %>%
    st_sf() %>%
    dplyr::select(-`<NA>`) %>%
    na.omit() %>%
    ungroup()

vars_net.long <- 
  gather(vars_net, Variable, value, -geometry, -uniqueID)

vars <- unique(vars_net.long$Variable)
mapList <- list()

for(i in vars){
  mapList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(vars_net.long, Variable == i), aes(fill=value), colour = NA) +
      scale_fill_viridis(name = "") +
      labs(title=i) +
      map.theme()}

do.call(grid.arrange,c(mapList, ncol = 3, top = "Risk Factors by Fishnet"))
```

## Feature Engineering

### NN

```{r create nn}
st_c <- st_coordinates
st_coid <- st_centroid

vars_net <-
  vars_net %>%
    mutate(
      abd_vehicle.nn =
        nn_function(st_c(st_coid(vars_net)), st_c(abd_vehicle),5),
      dan_building.nn =
        nn_function(st_c(st_coid(vars_net)), st_c(dan_building),5),
      graffiti.nn =
        nn_function(st_c(st_coid(vars_net)), st_c(graffiti),5),
      sanitation.nn =
        nn_function(st_c(st_coid(vars_net)), st_c(sanitation),5),
      st_light_out.nn =
        nn_function(st_c(st_coid(vars_net)), st_c(st_light_out),5),
      dan_park.nn =
        nn_function(st_c(st_coid(vars_net)), st_c(dan_park),5),
      onsite_liquor.nn =
        nn_function(st_c(st_coid(vars_net)), st_c(onsite_liquor),1),
      liquor_retail.nn =
        nn_function(st_c(st_coid(vars_net)), st_c(liquor_retail),1),
      po_station.nn = 
        nn_function(st_c(st_coid(vars_net)), st_c(po_station),1))

mean(vars_net$abd_vehicle.nn)
mean(vars_net$dan_building.nn)
mean(vars_net$graffiti.nn)
mean(vars_net$sanitation.nn)
mean(vars_net$st_light_out.nn)
mean(vars_net$dan_park.nn)
mean(vars_net$onsite_liquor.nn)
mean(vars_net$liquor_retail.nn)
mean(vars_net$po_station.nn)
```

### Centrality of Places

Take Center City as the center of Philadelphia.

```{r center}
cc_point <-
  filter(neigh, NAME == "CENTER_CITY") %>%
  dplyr::select(NAME, geometry) %>% 
  st_centroid()

vars_net$cc_distance =
  st_distance(st_centroid(vars_net),cc_point) %>%
  as.numeric()

ggplot() +
  geom_sf(data=vars_net, aes(fill=cc_distance)) +
  scale_fill_continuous_sequential(palette = "Viridis", trans = "reverse") +
  labs(title="Euclidean Distance to Center City") +
  map.theme() 
```

### Final Net

Join neighborhood and police district information.

```{r final net}
final_net <-
  left_join(crime_net, st_drop_geometry(vars_net), by="uniqueID") 

final_net <-
  st_centroid(final_net) %>%
    st_join(dplyr::select(neigh, name = NAME), by = "uniqueID") %>%
    st_join(dplyr::select(po_district, District = DISTRICT_), by = "uniqueID") %>%
      st_drop_geometry() %>%
      left_join(dplyr::select(final_net, geometry, uniqueID)) %>%
      st_sf() %>%
  na.omit()
```

### Spatial Process

```{r local moran}
## generates warnings from PROJ issues
## {spdep} to make polygon to neighborhoods... 
final_net.nb <- poly2nb(as_Spatial(final_net), queen=TRUE)
## ... and neighborhoods to list of weigths
final_net.weights <- nb2listw(final_net.nb, style="W", zero.policy=TRUE)

local_morans <- localmoran(final_net$count_felony, final_net.weights, zero.policy=TRUE) %>% 
  as.data.frame()

final_net.localMorans <- 
  cbind(local_morans, as.data.frame(final_net)) %>% 
  st_sf() %>%
  dplyr::select(Felony_Count = count_felony, 
                Local_Morans_I = Ii, 
                P_Value = `Pr(z != E(Ii))`) %>%
  mutate(Significant_Hotspots = ifelse(P_Value <= 0.001, 1, 0)) %>%
  gather(Variable, Value, -geometry)
  
## This is just for plotting
vars <- unique(final_net.localMorans$Variable)
varList <- list()

for(i in vars){
  varList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(final_net.localMorans, Variable == i), 
              aes(fill = Value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      map.theme(title_size = 14) + theme(legend.position="bottom")}

do.call(grid.arrange,c(varList, ncol = 4, top = "Local Morans I Statistics, Felony"))
```

```{r hot spot}
# generates warning from NN
final_net <- final_net %>% 
  mutate(felony.isSig = 
           ifelse(local_morans[,5] <= 0.001, 1, 0)) %>%
  mutate(felony.isSig.dist = 
           nn_function(st_c(st_coid(final_net)),
                       st_c(st_coid(filter(final_net, 
                                           felony.isSig == 1))), 
                       k = 1))

ggplot() +
      geom_sf(data = final_net, aes(fill = felony.isSig.dist), colour = NA) +
      scale_fill_viridis(name = "NN Distance") +
      labs(title="Distance to Highly Significant Felony Hotspots") +
      map.theme()
```

### Correlation Test

```{r correlation}
correlation.long <-
  st_drop_geometry(final_net) %>%
    dplyr::select(-uniqueID, -cvID, -cc_distance, -name, -District) %>%
    gather(Variable, Value, -count_felony)

correlation.cor <-
  correlation.long %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, count_felony, use = "complete.obs"))
    
ggplot(correlation.long, aes(Value, count_felony)) +
  geom_point(size = 0.1) +
  geom_text(data = correlation.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "orange", lwd = 0.7) +
  facet_wrap(~Variable, ncol = 3, scales = "free") +
  labs(title = "Felony Count as A Function of Risk Factors") +
  plot.theme()
```

## Modeling and Cross Validation

### Poison Regression

```{r poison regression}
ggplot(final_net, aes(count_felony)) + 
  geom_histogram(binwidth = 1) +
  geom_vline(aes(xintercept = mean(count_felony)),
             color = "orange", size = 0.7) +
  labs(title = "Distribution of Felony by Grid Cell") +
  plot.theme() +
  theme(plot.title = element_text(size = 10))
```

### Cross Validation

```{r cv}
reg.vars <- c("abd_vehicle.nn", "dan_building.nn", "graffiti.nn", "sanitation.nn", "st_light_out.nn",  "liquor_retail.nn", "cc_distance")

reg.ss.vars <- c("abd_vehicle.nn", "dan_building.nn", "graffiti.nn", "sanitation.nn", "st_light_out.nn",  "liquor_retail.nn", "cc_distance", "felony.isSig", "felony.isSig.dist")

reg.vars1 <- c("abd_vehicle.nn", "dan_building.nn", "graffiti.nn", "sanitation.nn", "st_light_out.nn",  "dan_park.nn","onsite_liquor.nn", "liquor_retail.nn", "po_station.nn", "cc_distance")

reg.ss.vars1 <- c("abd_vehicle.nn", "dan_building.nn", "graffiti.nn", "sanitation.nn", "st_light_out.nn",  "dan_park.nn","onsite_liquor.nn", "liquor_retail.nn", "po_station.nn", "cc_distance", "felony.isSig", "felony.isSig.dist")

## RUN REGRESSIONS
reg.cv <- crossValidate(
  dataset = final_net,
  id = "cvID",
  dependentVariable = "count_felony",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = cvID, count_felony, Prediction, geometry)

reg.ss.cv <- crossValidate(
  dataset = final_net,
  id = "cvID",
  dependentVariable = "count_felony",
  indVariables = reg.ss.vars) %>%
    dplyr::select(cvID = cvID, count_felony, Prediction, geometry)
  
reg.spatialCV <- crossValidate(
  dataset = final_net,
  id = "name",
  dependentVariable = "count_felony",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = name, count_felony, Prediction, geometry)

reg.ss.spatialCV <- crossValidate(
  dataset = final_net,
  id = "name",
  dependentVariable = "count_felony",
  indVariables = reg.ss.vars) %>%
    dplyr::select(cvID = name, count_felony, Prediction, geometry)

reg.spatialCV2 <- crossValidate(
  dataset = final_net,
  id = "District",
  dependentVariable = "count_felony",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = District, count_felony, Prediction, geometry)

reg.ss.spatialCV2 <- crossValidate(
  dataset = final_net,
  id = "District",
  dependentVariable = "count_felony",
  indVariables = reg.ss.vars) %>%
    dplyr::select(cvID = District, count_felony, Prediction, geometry)

reg.summary <- 
  rbind(
    mutate(reg.cv,           Error = Prediction - count_felony,
                             PE = (Prediction - count_felony)/count_felony,
                             Regression = "Random k-fold CV: Just Risk Factors"),
                             
    mutate(reg.ss.cv,        Error = Prediction - count_felony,
                             PE = (Prediction - count_felony)/count_felony,
                             Regression = "Random k-fold CV: Spatial Process"),
    
    mutate(reg.spatialCV,    Error = Prediction - count_felony,
                             PE = (Prediction - count_felony)/count_felony,
                             Regression = "Spatial LOGO-CV by neighborhoods: Just Risk Factors"),
                             
    mutate(reg.ss.spatialCV, Error = Prediction - count_felony,
                             PE = (Prediction - count_felony)/count_felony,
                             Regression = "Spatial LOGO-CV by neighborhoods: Spatial Process"),
    
    mutate(reg.spatialCV2, Error = Prediction - count_felony,
                              PE = (Prediction - count_felony)/count_felony,
                             Regression = "Spatial LOGO-CV by police districts: Just Risk Factors"),
    
    mutate(reg.ss.spatialCV2, Error = Prediction - count_felony,
                              PE = (Prediction - count_felony)/count_felony,
                             Regression = "Spatial LOGO-CV by police districts: Spatial Process")) %>%
    st_sf() 
```

```{r cv result}
reg.summary <- transform(reg.summary, Regression = factor(Regression, levels=c("Random k-fold CV: Just Risk Factors", "Spatial LOGO-CV by neighborhoods: Just Risk Factors", "Spatial LOGO-CV by police districts: Just Risk Factors", "Random k-fold CV: Spatial Process", "Spatial LOGO-CV by neighborhoods: Spatial Process", "Spatial LOGO-CV by police districts: Spatial Process")))

grid.arrange(
  reg.summary %>%
    ggplot() +
      geom_sf(aes(fill = Prediction), color = "transparent", lwd = 0.05) +
      geom_sf(data = phl_boundary, fill = "transparent", color = "black", lwd = 0.3) +
      facet_wrap(~Regression) +
      scale_fill_continuous_sequential(palette = "Reds 3", limits = c(0, 160)) +
      labs(title = "Predicted Felony Possessions by Regression") + 
      theme(
    plot.title = element_text(size = 10, hjust = 0.5), 
    plot.subtitle = element_text(size = 8), 
    plot.caption = element_text(size = 8,hjust = 1), 
    strip.text = element_text(size = 4), 
    axis.title = element_text(size = 0),
    axis.text = element_text(size = 0),
    legend.key.width = unit(0.2, "cm"),
    legend.title = element_text(size = 0), 
    legend.text = element_text(size = 4), 
    plot.margin = unit(c(0, 0, 0, 0),"cm"), 
    panel.background = element_rect(fill = "transparent"),
    panel.grid = element_line(color = "grey85", size = 0.1), 
    legend.position = "none"),

  filter(reg.summary, Regression == "Random k-fold CV: Just Risk Factors") %>%
    ggplot() +
      geom_sf(aes(fill = count_felony), color = "transparent", lwd = 0.05) +
      geom_sf(data = phl_boundary, fill = "transparent", color = "black", lwd = 0.5) +
      scale_fill_continuous_sequential(palette = "Reds 3", limits = c(0, 160)) +
      labs(title = "Felony") +
      theme(
    plot.title = element_text(size = 10, hjust = 0.5), 
    plot.subtitle = element_text(size = 8), 
    plot.caption = element_text(size = 8,hjust = 1), 
    strip.text = element_text(size = 8), 
    axis.title = element_text(size = 0),
    axis.text = element_text(size = 4),
    legend.key.width = unit(0.2, "cm"),
    legend.title = element_text(size = 0), 
    legend.text = element_text(size = 4), 
    plot.margin = unit(c(0, 0, 0, 0),"cm"), 
    panel.background = element_rect(fill = "transparent"),
    panel.grid = element_line(color = "grey85", size = 0.1), 
    legend.position = "right"), ncol = 2, widths = c(1.15, 1))
```

```{r comparing errors}
ggplot() +
    geom_sf(data = reg.summary, aes(fill = Error), color = "black", lwd = 0.05)+
    facet_wrap(~Regression, ncol = 3) +
    scale_fill_continuous_diverging(palette = "Blue-Red 3", trans = "reverse") +
    labs(title = "Predicted Felony Errors by Regression") +
    theme(
    plot.title = element_text(size = 10, hjust = 0.5), 
    plot.subtitle = element_text(size = 8), 
    plot.caption = element_text(size = 8,hjust = 1), 
    strip.text = element_text(size = 6), 
    axis.title = element_text(size = 0),
    axis.text = element_text(size = 4),
    legend.key.width = unit(0.2, "cm"),
    legend.title = element_text(size = 0), 
    legend.text = element_text(size = 4), 
    plot.margin = unit(c(0, 0, 0, 0),"cm"), 
    panel.background = element_rect(fill = "transparent"),
    panel.grid = element_line(color = "grey85", size = 0.1), 
    legend.position = "right")
```

```{r error plot}
reg.summary <- transform(reg.summary, Regression = factor(Regression, levels=c("Random k-fold CV: Just Risk Factors", "Random k-fold CV: Spatial Process", "Spatial LOGO-CV by neighborhoods: Just Risk Factors", "Spatial LOGO-CV by neighborhoods: Spatial Process", "Spatial LOGO-CV by police districts: Just Risk Factors", "Spatial LOGO-CV by police districts: Spatial Process")))

error_by_reg_and_fold <- 
  reg.summary %>%
    group_by(Regression, cvID) %>% 
    summarize(MAE = mean(abs(Error), na.rm = T),
              SD_MAE = mean(sd(abs(Error)), na.rm = T)) %>%
  ungroup()

error_by_reg_and_fold2 <- 
   reg.summary %>%
    group_by(Regression, cvID) %>%
    dplyr::filter(!is.infinite(PE)) %>% 
    summarize(MAPE = mean(abs(PE), na.rm = T),
              SD_MAPE = mean(sd(abs(PE)), na.rm = T)) %>%
  ungroup()

error_by_reg_and_fold %>%
  ggplot(aes(MAE)) + 
    geom_histogram(bins = 30, colour="black", fill = "#f78d8d", lwd = 0.5) +
    facet_wrap(~Regression, ncol = 2) +  
    geom_vline(xintercept = 0) + scale_x_continuous(breaks = seq(0, 8, by = 1)) + 
    labs(title="Distribution of MAE", subtitle = "k-fold cross validation vs. LOGO-CV",
         x="Mean Absolute Error", y="Count") +
  theme(
    plot.title = element_text(size = 10, hjust = 0.5), 
    plot.subtitle = element_text(size = 8), 
    plot.caption = element_text(size = 8,hjust = 1), 
    strip.text = element_text(size = 6), 
    axis.title = element_text(size = 0),
    axis.text = element_text(size = 6),
    legend.key.height = unit(0.2, "cm"),
    legend.title = element_text(size = 0), 
    legend.text = element_text(size = 6), 
    plot.margin = unit(c(0, 0, 0, 0),"cm"), 
    panel.background = element_rect(fill = "transparent"),
    panel.grid = element_line(color = "grey85", size = 0.1), 
    legend.position = "none")
```

```{r table}
error_by_reg_and_fold3 <-
  left_join(st_drop_geometry(error_by_reg_and_fold), st_drop_geometry(error_by_reg_and_fold2)) 

e.summary <- error_by_reg_and_fold3 %>% 
  group_by(Regression) %>% 
    summarize(Mean_MAE = round(mean(MAE), 2),
              SD_MAE = round(sd(MAE), 2))

kbl(format = 'html', escape = FALSE, e.summary, caption = "Mean Absolute Errors Across Different Models") %>%
  kable_styling(font_size = 10, full_width = F) %>%
  gsub("font-size: initial !important;", 
         "font-size: 10pt !important;", .)
```


```{r performance}
st_set_geometry(reg.summary, NULL) %>%
  group_by(Regression) %>%
    mutate(felony_decile = ntile(count_felony, 10)) %>%
  group_by(Regression, felony_decile) %>%
    summarize(meanObserved = mean(count_felony, na.rm=T),
              meanPrediction = mean(Prediction, na.rm=T)) %>%
    gather(Variable, Value, -Regression, -felony_decile) %>%          
    ggplot(aes(felony_decile, Value, shape = Variable)) + geom_path(aes(group = felony_decile), colour = "brown")  +
      geom_point(size = 2, colour = "#f78d8d", fill = "#f78d8d")+
      scale_shape_manual(values = c(2, 17)) +
      facet_wrap(~Regression) + xlim(0,10) +
      labs(title = "Predicted and Observed Narcotics Possession by Observed Narcotics Possession") +
    theme(
    plot.title = element_text(size = 10, hjust = 0.5), 
    plot.subtitle = element_text(size = 8), 
    plot.caption = element_text(size = 8,hjust = 1), 
    strip.text = element_text(size = 6), 
    axis.title = element_text(size = 0),
    axis.text = element_text(size = 6),
    legend.key.height = unit(0.2, "cm"),
    legend.title = element_text(size = 0), 
    legend.text = element_text(size = 6), 
    plot.margin = unit(c(0, 0, 0, 0),"cm"), 
    panel.background = element_rect(fill = "transparent"),
    panel.grid = element_line(color = "grey85", size = 0.1), 
    legend.position = "bottom")
```

## Generalizability by Race and Income Context

We don't use demographic features to build the model, but the projected felony incidents show socioeconomic bias in disadvantaged neighborhoods. This reveals the selection bias of intitial crime data set.

```{r get census data}
phl_mhhinc <- 
  get_acs(geography = "county", variables = "B19013_001", 
          year = 2022, state = "PA", county = "Philadelphia") %>% 
  dplyr::select(MedInc = estimate) %>% 
  pull() %>% 
  .[1]

tracts_22 <- 
  get_acs(geography = "tract", variables = c("B01003_001", "B01001A_001E", "B19013_001"), 
          year = 2022, state = "PA", county = "Philadelphia", geometry = T) %>%
  st_transform(crs = phl_crs)  %>% 
  dplyr::select(variable, estimate, GEOID) %>%
  spread(variable, estimate) %>%
  rename(TotalPop = B01003_001,
         NumberWhites = B01001A_001,
         MedInc = B19013_001) %>%
  mutate(percentWhite = NumberWhites / TotalPop,
         percentInc = MedInc / phl_mhhinc,
         Race_Context = ifelse(percentWhite > .5, "Majority_White", "Majority_Non_White"),
         Income_Context = case_when(
           percentInc <= .5 ~ "Low_Income",
           percentInc > .5 & percentInc <= 1.5 ~ "Moderate_Income",
           percentInc > 1.5 ~ "High_Income")) %>%
  .[neigh,]
  
tracts_22_map <- 
  tracts_22 %>% 
  dplyr::select(-TotalPop, -NumberWhites, -MedInc, -percentWhite, -percentInc) %>% 
  st_drop_geometry() %>% 
  gather(category, value, -GEOID) %>% 
  left_join(tracts_22) %>% 
  st_as_sf()
  

ggplot() + 
  geom_sf(data = tracts_22_map, aes(fill = value)) +
  scale_fill_discrete_sequential(palette = "Sunset") +
  facet_wrap(~category) +
  labs(title =  "Racial and Income Contexts in Philadelphia") +
  map.theme() 
```

```{r race and income}
reg.summary_race <- reg.summary %>% 
    st_centroid() %>%
    st_join(tracts_22) %>%
    na.omit() %>%
      st_drop_geometry() %>%
      group_by(Regression, Race_Context) %>%
      summarize(mean.Error = mean(Error, na.rm = T)) %>%
      spread(Race_Context, mean.Error)

reg.summary_inc <- reg.summary %>% 
    st_centroid() %>%
    st_join(tracts_22) %>%
    na.omit() %>%
      st_drop_geometry() %>%
      group_by(Regression, Income_Context) %>%
      summarize(mean.Error = mean(Error, na.rm = T)) %>%
      spread(Income_Context, mean.Error)
```

```{r race table}
kbl(format = 'html', escape = FALSE, reg.summary_race, caption = "Mean Error by Neighborhood Racial Context") %>%
      kable_styling(font_size = 10, full_width = F) %>%
      gsub("font-size: initial !important;", 
         "font-size: 10pt !important;", .)
```

```{r income table}
kbl(format = 'html', escape = FALSE, reg.summary_inc, caption = "Mean Error by Neighborhood Income Context") %>%
      kable_styling(font_size = 10, full_width = F) %>%
      gsub("font-size: initial !important;", 
         "font-size: 10pt !important;", .)
```

## Comparison with Kernel Density

```{r kernel density}
felony_ppp <- as.ppp(st_coordinates(felony_23), W = st_bbox(final_net))
felony_KD.1000 <- spatstat.explore::density.ppp(felony_ppp, 1500)

felony_KDE_sum <- as.data.frame(felony_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean)

kde_breaks <- classIntervals(felony_KDE_sum$value, 
                             n = 10, "fisher")

felony_KDE_sf <- felony_KDE_sum %>%
  mutate(label = "Kernel Density",
         Risk_Category = ntile(value, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 80 & Risk_Category <= 89 ~ "80% to 89%",
           Risk_Category >= 70 & Risk_Category <= 79 ~ "70% to 79%",
           Risk_Category >= 60 & Risk_Category <= 69 ~ "60% to 69%",
           Risk_Category >= 50 & Risk_Category <= 59 ~ "50% to 59%",
           Risk_Category >= 40 & Risk_Category <= 49 ~ "40% to 49%",
           Risk_Category >= 30 & Risk_Category <= 39 ~ "30% to 39%",
           Risk_Category >= 20 & Risk_Category <= 29 ~ "20% to 29%",
           Risk_Category >= 10 & Risk_Category <= 19 ~ "10% to 19%",
           Risk_Category >= 1 & Risk_Category <= 9 ~ "1% to 9%")) %>%
  cbind(
    aggregate(
      dplyr::select(felony_23) %>% mutate(felonyCount = 1), ., sum) %>%
    mutate(felonyCount = replace_na(felonyCount, 0))) %>%
  dplyr::select(label, Risk_Category, felonyCount)


felony_risk_sf <-
  filter(reg.summary, Regression == "Random k-fold CV: Spatial Process") %>%
  mutate(label = "Risk Predictions",
         Risk_Category = ntile(Prediction, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 80 & Risk_Category <= 89 ~ "80% to 89%",
           Risk_Category >= 70 & Risk_Category <= 79 ~ "70% to 79%",
           Risk_Category >= 60 & Risk_Category <= 69 ~ "60% to 69%",
           Risk_Category >= 50 & Risk_Category <= 59 ~ "50% to 59%",
           Risk_Category >= 40 & Risk_Category <= 49 ~ "40% to 49%",
           Risk_Category >= 30 & Risk_Category <= 39 ~ "30% to 39%",
           Risk_Category >= 20 & Risk_Category <= 29 ~ "20% to 29%",
           Risk_Category >= 10 & Risk_Category <= 19 ~ "10% to 19%",
           Risk_Category >= 1 & Risk_Category <= 9 ~ "1% to 9%")) %>%
  cbind(
    aggregate(
      dplyr::select(felony_23) %>% mutate(felonyCount = 1), ., sum) %>%
      mutate(felonyCount = replace_na(felonyCount, 0))) %>%
  dplyr::select(label,Risk_Category, felonyCount)

felony_actual_sf <- felony_23 %>% 
  dplyr::select() %>% 
  mutate(felonyCount23 = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(felonyCount23 = replace_na(felonyCount23, 0),
         label = "Actual Value",
         Risk_Category = ntile(felonyCount23, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 80 & Risk_Category <= 89 ~ "80% to 89%",
           Risk_Category >= 70 & Risk_Category <= 79 ~ "70% to 79%",
           Risk_Category >= 60 & Risk_Category <= 69 ~ "60% to 69%",
           Risk_Category >= 50 & Risk_Category <= 59 ~ "50% to 59%",
           Risk_Category >= 40 & Risk_Category <= 49 ~ "40% to 49%",
           Risk_Category >= 30 & Risk_Category <= 39 ~ "30% to 39%",
           Risk_Category >= 20 & Risk_Category <= 29 ~ "20% to 29%",
           Risk_Category >= 10 & Risk_Category <= 19 ~ "10% to 19%",
           Risk_Category >= 1 & Risk_Category <= 9 ~ "1% to 9%")) %>% 
  mutate(felonyCount = felonyCount23,
         geometry = geometry)
felony_actual_sf <- felony_actual_sf[, -c(1, 2)]
  

rbind(felony_KDE_sf, felony_risk_sf, felony_actual_sf) %>%
  na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category, -geometry) %>%
  ggplot() +
    geom_sf(aes(fill = Risk_Category), colour = NA) +
    geom_sf(data = sample_n(felony_23, 3000), size = .3, colour = "black") +
    facet_wrap(~label, ) +
    scale_fill_viridis(discrete = TRUE) +
    labs(title="Comparison of Kernel Density and Risk Predictions",
         subtitle="2022 Felony risk predictions; 2023 Felony") +
    map.theme()
```

```{r kde comparison table}
rbind(felony_KDE_sf, felony_risk_sf, felony_actual_sf) %>%
  st_drop_geometry() %>%
  na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category) %>%
  group_by(label, Risk_Category) %>%
  summarize(countFelony = sum(Value)) %>%
  ungroup() %>%
  group_by(label) %>%
  mutate(Pcnt_of_test_set_crimes = countFelony / sum(countFelony)) %>%
    ggplot(aes(Risk_Category,Pcnt_of_test_set_crimes)) +
      geom_bar(aes(fill=label), position="dodge", stat="identity") +
      scale_fill_viridis(discrete = TRUE, name = "Model") +
      labs(title = "Risk prediction vs. Kernel density, 2023 Felony",
           y = "% of Test Set Felony (per model)",
           x = "Risk Category") +
  theme_bw() +
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

